### 2014-02-08 arfi_methods.tex
Stephen's comments

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe both the on and off-axis response of the tissue to the pushing pulses.
Specifically, four lines were dedicated to tracking the on-axis displacement,
with all 4 beams located inside the beamwidth of the pushing pulses such that
the tracking beam spacing was 0.17 mm.

3D volumetric imaging data was acquired using mechanical rotation of the ER7B
transducer between sequential imaging frames, sweeping across the lateral
extent of the prostate.  This rotation setup utilized a CIVCO
Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA USA)
that allows for complete 6-axis degrees of freedom for manual positioning of
the transducer to sweep through the entire prostate during imaging.  Sequential
imaging frames were $\sim$1\degree apart, depending on the absolute size of the
prostate.  A custom optical angular feedback transduction circuit utilizing a
reflective linear strip with 500 lines-per-inch (LPI) resolution (US Digital,
Vancouver, WA, USA) was coupled with a 141 oz-in torque stepper motor with
planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation, Anaheim, CA,
USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate spatial
localization of the imaging frames in the 3D dataset.

@@

Displacement estimation was performed using a phase-shift estimator on the
beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}. The ARFI
data were then normalized as a function of depth to account for attenuation and
focal gain effects.  This normalization was performed using a displacement
profile measured in a homogeneous tissue-mimicking phantom, which was applied
to all displacements in the entire data set at each time step, and then
low-pass filtered with a cutoff-frequency of 0.8 mm$^{-1}$.

The normalized displacement data were scan-converted using a 3D Delaunay
triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
Natick, MA, USA).  The scan-converted ARFI imaging data had a final isotropic
voxel size of 0.15 $\times$ 0.15 $\times$ 0.15 mm$^3$.

>>>>

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe the response of the tissue to the pushing pulses.  Specifically, four
lines were dedicated to tracking the on-axis displacement, with all 4 beams
located inside the beamwidth of the pushing pulses such that the tracking beam
spacing was 0.17 mm.  The tracking beams were focused at 60 mm with F/3
transmit and F/0.5 receive focal configurations.

3D volumetric imaging data was acquired using mechanical rotation of the ER7B
transducer between sequential imaging frames, sweeping across the lateral
extent of the prostate.  This rotation setup utilized a CIVCO
Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA USA)
that allows for complete 6-axis degrees of freedom for manual positioning of
the transducer to sweep through the entire prostate during imaging.  Sequential
imaging frames were $\sim$1\degree apart, depending on the absolute size of the
prostate.  A custom optical angular feedback transduction circuit utilizing a
reflective linear strip with 212 lines-per-inch (LPI) resolution (US Digital,
Vancouver, WA, USA) with a QSB-S Quadrature-to-USB adapter to achieve 9 line /
degree resolution.  Rotation was performed with a 141 oz-in torque stepper
motor with planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation,
Anaheim, CA, USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate
spatial localization of the imaging frames in the 3D dataset.

@@

Displacement estimation was performed using a phase-shift estimator on the
beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}. The ARFI
data were then normalized as a function of depth to account for attenuation and
focal gain effects.  This normalization was performed using a displacement
profile measured in a homogeneous tissue-mimicking phantom that was low-pass
filtered with a cutoff-frequency of 0.8 mm$^{-1}$.  This filtered normalization
profile was applied to all displacements in the entire data set at each time
step.

The normalized displacement data were scan-converted using a 2D Delaunay
triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
Natick, MA, USA) for each lateral location in the dataset.  The scan-converted
ARFI imaging data had a final isotropic voxel size of 0.15 $\times$ 0.15
$\times$ 0.15 mm$^3$.

### 2014-02-08 conclusions.tex
Stephen's comments

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations between estimated organ
volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
$\pm$ \ARFImrTotalLatLatStdPct\%), while differences in the other dimensions
were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
bias between the imaging modalities, but significant variability case-to-case
(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
volume differences were, again, strongly attributed to over-estimation of the
lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
apex-to-base dimension (\ARFImrCentralAntPostMeanPct~$\pm$
\ARFImrCentralAntPostStdPct\%).  Strong variability in CG volumes is believed
to be related to the extent of BPH for select cases.  Overall, ARFI imaging of
the prostate yielded prostate volumes and dimensions that were correlated with
MR T2WI estimates, with differences in the lateral and apex-to-base dimensions.
Poor image contrast between the prostate and peri-prostatic fat may be one
cause of these differences.  ARFI imaging is a promising low-cost, real-time
imaging modality that can compliment MR imaging for diagnosis, treatment
planning and management of PCa.

>>>>

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations between estimated organ
volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
$\pm$ \ARFImrTotalLatLatStdPct\%), while differences in the other dimensions
were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\%).  The central gland volumes of ARFI and MR images
were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
bias between the imaging modalities, but significant variability case-to-case
(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%).  Central gland
volume differences were, again, strongly attributed to over-estimation of the
lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
apex-to-base dimension (\ARFImrCentralAntPostMeanPct~$\pm$
\ARFImrCentralAntPostStdPct\%).  Strong variability in CG volumes is believed
to be related to the extent of BPH for select cases.  Overall, ARFI imaging of
the prostate yielded prostate volumes and dimensions that were correlated with
MR T2WI estimates, with differences in the lateral and apex-to-base dimensions.
Poor image contrast between the prostate and peri-prostatic fat may be one
cause of these differences.  ARFI imaging is a promising low-cost, real-time
imaging modality that can compliment MR imaging for diagnosis, treatment
planning and management of PCa.

### 2014-02-08 image_proc_methods.tex
Stephen's comments

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP (\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base-to-apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented by an experienced radiologist
using the smooth polygon tool in ITK-SNAP
(\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for
the PZ, CG and AFS. The gland was segmented from base-to-apex.  The base was
identified below the bladder, and subsequent images were segmented until the
last slice with visible prostatic tissue was identified caudally. The CG,
urethra, PZ and AFS were segmented independently according to their
well-established anatomical characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

### 2014-02-08 mr_background.tex
Stephen's comments

\subsubsection{T2-Weighted Imaging and Prostate Anatomy}
T2WI sequences are crucial components of prostate MR imaging.  T2WI is
particularly useful in prostate analysis due to its excellent soft tissue
contrast resolution, which can be maximized by using thin sections of 3--4 mm
and a small field of view of approximately 14 cm.~\cite{Gupta2013,Bonekamp2011}
T2 sequences are the most helpful for tumor localization, as they can clearly
show overall prostate morphology, internal structures and prostatic
margins.~\cite{Gupta2013}

>>>>

\subsubsection{T2-Weighted Imaging and Prostate Anatomy}
T2WI sequences are crucial components of prostate MR imaging.  T2WI is
particularly useful in prostate analysis due to its excellent soft tissue
contrast resolution, which can be maximized by using thin sections of 3--4 mm
and a small field of view of approximately 14
cm$^2$.~\cite{Gupta2013,Bonekamp2011} T2 sequences are the most helpful for
tumor localization, as they can clearly show overall prostate morphology,
internal structures and prostatic margins.~\cite{Gupta2013}

### 2014-02-08 tab_mr_arfi_volumes.tex
updated figures for correlation analyses and removed image comparisons to path axes
Stephen's comments

\begin{table}
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Zone} & {\bf ARFI Total} \\ 
& {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline
\input{data/vol_data}
\hline
\end{tabular}
\label{tab:mr_arfi_volumes}
\end{table}

>>>>

\begin{table}[h!]
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Gland} & {\bf ARFI Total} \\ 
& {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline
\input{data/vol_data}
\hline
\end{tabular}
\label{tab:mr_arfi_volumes}
\end{table}
